---
title: |
  | Exporing Discretionary Accruals: 
  | A Stub Project^[I acknowledge financial support from Deutsche Forschungsgemeinschaft-Project-ID 403041268 - TRR 266.] \vspace{1cm}

author: |
  | Joachim Gassen 
  | TRR 266 Accounting for Transparency
  | Humboldt-Universität zu Berlin 
  | gassen@wiwi.hu-berlin.de
  |
  | Ulrich Atz
  | New York University
  | Stern School of Business
  | ua6@stern.nyu.edu

date: |
  | `r loc <- Sys.getlocale(category = "LC_TIME"); Sys.setlocale("LC_TIME", "C"); fdate <- format(Sys.time(), '%B %e, %Y'); Sys.setlocale("LC_TIME", loc); fdate` \vspace{1cm}
  
abstract: |
  | The Open Science movement promotes the accessibility and reusability of research. This repository has the objective to help researchers establishing such an collaboration-oriented workflow. It uses a toy project on discretionary accruals for demonstration. 
  | \vspace{6cm}

bibliography: references.bib
biblio-style: apsr

output:
  pdf_document: 
    number_sections: true
  toc: no
fig_caption: yes
fontsize: 11pt
ident: yes

always_allow_html: yes

header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
library(ExPanDaR)
library(kableExtra)
library(patchwork)
opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
opts_chunk$set(out.width = '100%', dpi=300)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache = FALSE)
```

```{r load_results, include=FALSE}
load("../output/results.rda")
```

\pagebreak

# Introduction

This is not a paper but a stub that is part of a [project template repository](https://github.com/trr266/treat). We developed this repository to provide a prototype for a reproducible and collaborative workflow. Several authors have discussed advantages of such workflows (@W_2017, @GGR_2018, @CFM_2019) and many scholars across fields have voiced the need for increased reproducibility in science (e.g., @I_2005, @GL_2014, @DPR_2017).

# Discretionary Accruals

To demonstrate our workflow, we explore discretionary accruals across the U.S. We calculate modified Jones and Dechow and Dichev type accruals and show their distributional properties. The main purpose of all this, however, is to provide a toy use case for [our project template directory](https://github.com/trr266/project_template) that contains all the code to obtain the data, run the analysis and prepare a paper as well as a presentation.

\ref{tab:descriptives} presents our data that is based on a simple WRDS pull of Compustat data with financial firms (SIC 6XXX) excluded. We require data to calculate all variables and this drastically reduces the sample size. Modified Jones discretionary accruals are calculated loosely based on @HN_2007 and Dechow and Dichev discretionary accruals are calculated based on (big surprise) @DD_2002. As you will see from \ref{fig:boxplot_smp}, discretionary accruals are very noisy constructs, even after limiting the sample to observations with complete data and winsorizing all data to the top and bottom percentile for each year. Figure \ref{fig:scatter_dd_lnta} shows a very prominent heteroscedasticity of discretionary accruals with regards to size. While researchers have tried to address this problem, the distributional properties of these constructs significantly complicate the interpretation of discretionary accrual-related findings. Especially in high powered settings, the measurement error, being highly correlated with size, will tend to load on variables that are unrelated to the underlying economic construct but correlated with size. \ref{tab:corr} shows some correlations and \ref{tab:reg} shows some completely pointless regressions.

# Discretionary Accruals, revisited

The shortcomings of the traditional accrual models are well documented, for example, @ball2013 questions whether the standard models for discretionary accruals measure what they promise to do. Recently, scholars have proposed a Bayesian methodology [@breuer2019a] to address some aspects (such as extreme estimates and false positives) of this criticism. The core idea, "borrowing strength" as statisticians call it, i.e. shrinking extreme parameters toward a more reasonable grand mean, may also be achieved be classic frequentist methods such as hierarchical models (also known as mulitlevel or mixed model). They come at a computational cost that are orders of magnitude cheaper.

Figure \ref{fig_boxplot_coef} shows how the two model approaches compare in their size of coefficients for each industry-year. As discussed, the larger outliers in the traditional approach (one regression for each group) are subdued in the hierarchical approach.

This desirable feature comes at a cost. We are only estimating a dozen parameters instead of more than a thousand, so that the in-sample predictions are perhaps wores. \ref{fig:boxplot_mjh} and \ref{fig:scatter_mjh} show how the modified Jones model agrees with the hierarchical Jones model in their point estimates for discretionary accruals. Contrary to expectation, the hierarchical model exhibits larger extreme values. We speculate this could also be due to a convergence issue of the model. The R package we employed is `lme4` [@bates2015].

Firm-level estimates were often not available because they require a sufficient number of observations per firm. With the hierarchical model we mitigate this issue. \ref{fig_boxplot_coef2} shows how the firm-level model  compares in their size of coefficients to the classic modified Jones model. Again, the larger outliers in the traditional approach (one regression for each group) are subdued in the hierarchical approach. The point estimates are again larger than before (\ref{fig:boxplot_mjh2} and \ref{fig:scatter_mjh2}). 

Finally, it may be possible to combine the benefits of a hierarchical approach with the benefits of estimating many parameters in a Bayesian model that specifies weak prior information. We outline this approach in the code, but do not present the results as they require further fine-tuning to be interpretable. 

```{r BoxPlot, echo = FALSE, fig.align="center", fig.cap="\\label{fig:boxplot_smp}Distribution of Discretionary Accruals over Time"}
fig_boxplot_smp
```

```{r Scatter, echo = FALSE, fig.align="center", fig.cap="\\label{fig:scatter_dd_lnta}Dechow and Dichev DA and Firm Size"}
fig_scatter_dd_lnta
```

```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:coef_mjh} Regression coefficients for the modified Jones model and its hierarchical equivalent (industry-year)"}
fig_boxplot_coef
```

```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:boxplot_mjh}Distribution of Discretionary Accruals over Time, revisited"}
fig_boxplot_hier
```

```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:scatter_mjh} Correlation between the modified Jones model and its hierarchical equivalent"}
fig_scatter_md_mjh
```


```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:coef_mjh2} Regression coefficients for the modified Jones model and its hierarchical equivalent (firm-level)"}
fig_boxplot_coef2
```

```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:boxplot_mjh2}Distribution of Discretionary Accruals over Time, revisited (firm-level)"}
fig_boxplot_hier2
```

```{r , echo = FALSE, fig.align="center", fig.cap="\\label{fig:scatter_mjh2} Correlation between the modified Jones model and its hierarchical equivalent (firm-level"}
fig_scatter_md_mjh2
```




```{r tab_descriptives_sample, results="asis"}
desc_rnames <- paste0("\\textit{", var_names$label, "}")
t <- tab_desc_stat
rownames(t$df) <- desc_rnames
names(t$df)[c(5,7)] <- c("25 \\%", "75 \\%")
kable(t$df, digits = c(0, 3, 3, 3, 3, 3, 3, 3), format = "latex",
      caption = "\\label{tab:descriptives}Descriptive Statistics",
      format.args = list(decimal.mark = ".", big.mark = ",", 
                         scientific=FALSE),
      booktabs = TRUE, escape = FALSE, linesep = "") -> kab_latex

lat_tab <- unlist(strsplit(kab_latex, "\n"))
lat_tab[6] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
lat_tab[8] <- "\\hline\\\\[-1.8ex]"
lat_tab[length(lat_tab) - 2] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"

latex_tab <- c(
  "\\begin{table}[!htbp]",  
  lat_tab[3],
  "\\begin{adjustbox}{width=\\textwidth}",
  "\\begin{threeparttable}",
  lat_tab[5:(length(lat_tab) - 1)],
  "\\begin{tablenotes}[flushleft]",
  "\\setlength{\\labelsep}{0pt}",
  "\\footnotetext",
  sprintf(
    paste(
      "\\item Note: The data is obtained from the Compustat U.S. as provided",
      "by WRDS. The sample covers the period %d to %d and %s unique firms."
    ),
    min(smp_da$fyear), max(smp_da$fyear), 
    format(length(unique(smp_da$gvkey)), big.mark = ",")
  ),
  "\\end{tablenotes}",
  "\\end{threeparttable}",
  "\\end{adjustbox}",
  "\\end{table}"
)
cat(paste(latex_tab, collapse = "\n"))  
```

```{r CorrTable, results="asis"}
tab <- tab_corr
lat_tab <- unlist(strsplit(tab$kable_ret, "\n"))
lat_tab[7:13] <- str_replace(
  lat_tab[7:13], fixed(var_names$var_name), var_names$label
)
lat_tab[4] <- "\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]"
lat_tab[6] <- "\\hline\\\\[-1.8ex]"
lat_tab[14] <- "\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]"
lat_tab[16] <- "\\begin{tablenotes}[flushleft]"

latex_tab <- c(
  "\\begin{table}[!htbp]",  
  "\\centering",
  "\\caption{\\label{tab:corr}Correlations}",
  lat_tab,
  "\\end{table}"
)
cat(paste(latex_tab, collapse = "\n"))
```

```{r RegTable, results="asis"}
tab <- tab_regression$table
tab[12] <- paste(
  "\\\\[-1.8ex] &", var_names$label[1], 
  "&", var_names$label[2], "\\\\"
)
esc_var_names <- str_replace_all(var_names$var_name, fixed ("_"), "\\_")
reg_names <- tibble(
  var_name = esc_var_names[c(3, 5:7)],
  label = var_names$label[c(3, 5:7)]
)
for (i in seq(15, 24, by = 3)) {
  pos <- (i-12)/3
  tab[i] <- str_replace(
    tab[i], fixed(reg_names$var_name[pos]), reg_names$label[pos]
  )
}
latex_tab <- c(
  "\\begin{table}[!htbp]",  
  "\\caption{\\label{tab:reg}Regressions}",
  "\\centering",
  "\\begin{tabular}{lcc}",
  tab[8:36],
  "\\end{tabular}",
  "\\end{table}"
)
cat(paste(latex_tab, collapse = "\n"))
```

# Conclusion

Isn't that wonderful? Discretionary accruals rock but what rocks even more is open science and a collaborative workflow. Clone or fork this repository to kickstart your own projects. If you do not like R, consider contributing code in your favorite statistical programming language to the repo. Thanks for reading and enjoy!

\pagebreak

```{=tex}
\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}
```
# References {.unnumbered}

```{=tex}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
```
\noindent
